\chapter{Data}

\section{Datasets}
This section details the datasets and technical indicators used for the machine learning algorithms to predict the next day's classification of BTC, ETH, and SOL prices. The datasets include historic prices, Google Trends, Wikipedia page views, and top 100 Reddit posts for sentiment analysis and topic modeling. Additionally, several technical indicators are engineered from the main price data tables.

\subsection{Historic Prices}
For BTC, ETH, and SOL, we collected historic prices with the following attributes:
\begin{itemize}
    \item \textbf{Price:} The closing price of the cryptocurrency.
    \item \textbf{Open:} The opening price.
    \item \textbf{High:} The highest price during the trading period.
    \item \textbf{Low:} The lowest price during the trading period.
    \item \textbf{Close:} The closing price.
    \item \textbf{Volume:} The trading volume.
    \item \textbf{Pct Change:} The percentage change in price.
\end{itemize}

\subsection{Google Trends}
Google Trends data is collected for BTC, ETH, and SOL to understand the relative search interest over time. This data helps gauge public interest and potential market movements.

\subsection{Wikipedia Page Views}
Wikipedia page views for BTC, ETH, and SOL are used to measure the general public's interest in these cryptocurrencies. This data serves as a proxy for market sentiment and public awareness.

\subsection{Reddit Posts}
Top 100 Reddit posts per respective cryptocurrency subreddit (BTC, ETH, SOL) are collected. The data includes:
\begin{itemize}
    \item \textbf{Title:} The title of the post.
    \item \textbf{Description:} A brief description of the post.
    \item \textbf{Date Posted:} The date the post was published.
    \item \textbf{Number of Likes:} The number of likes the post received.
    \item \textbf{Number of Comments:} The number of comments on the post.
    \item \textbf{Overall Score:} The overall score of the post.
\end{itemize}
This data will be used for sentiment analysis and topic modeling to gain insights into community sentiment and trending topics.

\subsection{TVL (Total Value Locked)}
Total Value Locked (TVL) represents the total capital held within a specific protocol, encompassing all the assets staked, loaned, or otherwise utilized. This metric provides insightful analytics on market confidence and the overall health of blockchain ecosystems. Higher TVL figures often correlate with increased media coverage and user interest, enhancing the platform's marketing appeal and perceived stability.

\subsection{Gold, S\&P 500, and VIX Historic Data}
\begin{itemize}
    \item \textbf{Gold Data:} Includes historical data on gold prices, including the opening, high, low, and closing prices. Gold is often considered a safe-haven investment.
    \item \textbf{S\&P 500 (SNP) Data:} Includes historical data for the S\&P 500 index, covering opening, high, low, and closing prices, as well as trading volume. It is a key indicator of the U.S. stock market's overall performance.
    \item \textbf{VIX Data:} The VIX measures market expectations of near-term volatility conveyed by S\&P 500 stock index option prices. It reflects investor uncertainty and market sentiment.
\end{itemize}

\subsection{Technical Indicators and Trading Signals}
This project includes various technical analysis indicators calculated using pandas. Below are the formulas used for each indicator, formatted for clarity.

\small
\begin{longtable}{@{}p{3.5cm}p{6cm}X@{}}
\toprule
\textbf{Indicator} & \textbf{Formula} & \textbf{Trading Signal Explanation} \\ \midrule
\endhead

10-Day Moving Average (10D MA) & $\text{10D MA} = \frac{1}{10} \sum_{i=0}^{9} \text{Price}_{t-i}$ & A price above the 10D MA suggests upward momentum, signaling a potential buy opportunity. \\ \midrule
30-Day Moving Average (30D MA) & $\text{30D MA} = \frac{1}{30} \sum_{i=0}^{29} \text{Price}_{t-i}$ & When the price is above the 30D MA, it indicates a medium-term upward trend, suggesting a buy signal. \\ \midrule
Stochastic Oscillator \%K & $\%K = 100 \times \frac{\text{Close} - \text{Low}_{14}}{\text{High}_{14} - \text{Low}_{14}}$ & An upward \%K signal indicates buying pressure, suggesting a bullish trend. \\ \midrule
Stochastic Oscillator \%D & $\%D = \frac{1}{3} (\%K_{t} + \%K_{t-1} + \%K_{t-2})$ & An upward \%D signal shows continued buying interest, confirming the bullish trend. \\ \midrule
Relative Strength Index (RSI) & $\text{RSI} = 100 - \frac{100}{1 + \frac{\text{Average Gain}}{\text{Average Loss}}}$ & RSI above 70 indicates overbought conditions (sell signal), below 30 indicates oversold (buy signal). \\ \midrule
Momentum & $\text{Momentum} = \text{Price}_{t} - \text{Price}_{t-n}$ & Positive momentum indicates a strong upward trend, suggesting a buy signal. \\ \bottomrule
\caption{Technical Indicators and Formulas (Part 1)}
\end{longtable}

\begin{longtable}{@{}p{3.5cm}p{6cm}X@{}}
\toprule
\textbf{Indicator} & \textbf{Formula} & \textbf{Trading Signal Explanation} \\ \midrule
\endhead

MACD & $\text{MACD} = \text{12D EMA} - \text{26D EMA}$ & An upward MACD signal indicates bullish momentum, suggesting a buy opportunity. \\ \midrule
Commodity Channel Index (CCI) & $\text{CCI} = \frac{\text{Price} - \text{MA}}{0.015 \times \text{MAD}}$ & CCI above 100 indicates overbought (sell), below -100 indicates oversold (buy). \\ \midrule
Bollinger Bands & $\text{Upper Band} = \text{MA} + 2 \times \text{STD}$, $\text{Lower Band} = \text{MA} - 2 \times \text{STD}$ & A price above the upper band suggests overbought conditions (sell), below the lower band suggests oversold (buy). \\ \midrule
Fibonacci Retracement & Levels at 23.6\%, 38.2\%, 50\%, 61.8\%, and 100\% & Used to identify potential reversal levels in the price movement. \\ \midrule
Average True Range (ATR) & $\text{ATR} = \frac{1}{n} \sum_{i=0}^{n-1} \text{TR}_{i}$, $\text{TR} = \max(\text{High} - \text{Low}, |\text{High} - \text{Close}_{prev}|, |\text{Low} - \text{Close}_{prev}|)$ & Higher ATR values indicate higher market volatility. \\ 
\bottomrule
\caption{Technical Indicators and Formulas (Part 2)}
\end{longtable}
\normalsize

\section{Final Dataset Overview}
The final dataset combines historical prices, Google Trends, Wikipedia page views, Reddit posts, TVL, and various technical indicators. These diverse data sources provide a comprehensive view of market dynamics and sentiment, enhancing the predictive power of the models.


